\documentclass[a4paper]{scrartcl}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{graphicx}
\usepackage{lastpage}
\usepackage{pgf}
\usepackage{wrapfig}
\usepackage{fancyvrb}
\usepackage{fancyhdr}
\pagestyle{fancy}

% Create header and footer
\headheight 27pt
\pagestyle{fancyplain}
\lhead{\footnotesize{Network Programming, ID1212}}
\chead{\footnotesize{Name of the Report}}
\rhead{}
\lfoot{}
\cfoot{\thepage\ (\pageref{LastPage})}
\rfoot{}

% Create title page
\title{Name of the Report}
\subtitle{Network Programming, ID1212}
\author{Your Name and Email Address}
\date{Date}

\begin{document}

\maketitle

\section*{Tips for Report Writing}
\textbf{REMOVE THIS SECTION BEFORE SUBMITTING THE REPORT.}\\

\noindent \textit{The target audience has exactly the same skills as the author, except they do not know anything at all about the specific program described in the report.} \\

Consider the following:

\begin{itemize}
  \item \textbf{The report must be \textit{centered around the requirements}. Which are they (Introduction), how did you work to meet them (Method), what is the solution that meets them (Result), and how can you be sure they are met (Discussion). This is the IMRaD method.}

  \item \textbf{The report must show that you have done the work yourself and that you have understood what you have done. Both of these goals are met by carefully explaining the source code.}

  \item Is spelling and grammar correct? Is spoken language avoided?

  \item Does the report have a good structure with sections, subsections and paragraphs?

  \item Is the solution clearly explained? Will the reader understand the program? What would you yourself want to know if you read about the program, is that included in the report?

  \item Is the solution analyzed and evaluated? Are important properties of the program explained? Should there have been more extensive evaluation?

  \item Is the text clarified with images and/or other figures, and with links to the code in your Git repository? Remember that all figures (images, tables, graphs, code listings, etc) shall be numbered and have a short explaining text.
\end{itemize}

\section{Introduction}

\textbf{This section tells \textit{what} are you going to do.} \\

\noindent Explain the task and the requirements on the solution. If you wrote the program together with other students, list their names and email addresses here.

\section{Literature Study}

This section must prove that you collected sufficient knowledge before starting development, instead of just hacking away without knowing how to complete the task. State what you have read and briefly summarize what you have learned.

\section{Method}

\textbf{This section tells \textit{how} you solved the task.} \\

\noindent Explain how you worked when solving the task and how you evaluated that your solution met the requirements. Mention development tools and IDEs you used. \textit{Do not explain your solution and do not refer to code}, that belongs to the \textit{Result} section.

\section{Result}

\textbf{This section explains \textit{the result} of what you did.} \\

\noindent Present your solution and prove that it meets the requirements listed in the homework specification. Make sure to  \textit{clearly explain how you met each requirement}. You can, for example, write a separate subsection for each requirement, where you mention the requirement and explain how you met it. You can prove that a requirement is met by, for example, explaining source code related to the requirement. If so, include a link to that particular code in the Git repository, or clearly tell which method and class you intend. You can also prove that a requirement is met by including, and explaining, screenshots of the user interface.

Also, prove that you participated in writing the program, and that you understand it in detail. You prove this by explaining essential parts of the source code.

\pagebreak

\section{Discussion}

\textbf{This section \textit{analysis} the result presented in the previous section.} \\

\noindent Summarize the requirements and \textit{clearly state which of them you have met}. This might seem like a meaningless repetition of the Result section, but it is not. Here, you shall not prove that the requirements are met, just briefly state what you achieved, that is, which requirements where met. This summary is very useful when assessing your report. Also, explain what lessons you learned, what problems you faced, and how the problems were solved. Should something have been done differently?

\section{Comments About the Course}

Any comment(s) related to this course offering or to coming offerings is much appreciated. \textit{Please also tell approximately how much time you spent on the assignment}, including lectures and exercises. This is of great help for improving the course.

\end{document}
